Stateless Model Checking with Data-Race Preemption Points

### 2015-11-18 pldi-16/evaluation.tex
implement brandon's comments, and trim a bit; back onto 12th page, oops
trimtrimtrim

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone in other model checkers.
For example, a single-state-space tool could use data-race candidates during immediately subsequent interleavings, essentially changing the state space on the fly.
Likewise, a message-passing-only tool could employ Iterative Deepening despite data races being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also sought to evaluate each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}

@@

\subsection{Finding nondeterministic data-race candidates}
\label{sec:eval-falseneg}

@@

>>>>

Although \quicksand~presents Iterative Deepening and data-race PPs as interconnected techniques, they each could theoretically be employed alone as well. %in other model checkers.
For example, a single-state-space tool could use data-race PPs during immediately subsequent interleavings,
%essentially
changing the state space on the fly.
Likewise, a message-passing-only tool could use Iterative Deepening despite a concurrency model lacking data races. % being absent from its concurrency model.
Hence, though many of our experiments compare \quicksand~to the state-of-the-art as a whole,
we also evaluated each technique individually.
Our evaluation answers the following questions:
\begin{enumerate}

@@

\subsection{Finding nondeterministic data-race candidates}
\label{sec:eval-falseneg}
%For the [NONDET] expt, you don't need to additionally count how many DRs
%needed to be found in a DR-PP state space, because when EXPLORE_BACKWARDS=0,
%you'll need to preempt on the DR PP to find the new DR anyway, hence they will
%be nondet. Unless you want to separately count the subset of nondet drs that
%are also dr-pp-ss-only (the reviewers may ask for this before cam-ready?).

@@

%When proving that [NONDET] dr buges exist, make sure to mention that, although
%their frequency varies depending on test case (mx test: almost none; bct:
%almost all), they are still PRESENT in all (or almost all) test cases, meaning
%it is not just a matter of writing better test cases.

### 2015-11-18 pldi-16/intro.tex
trimtrimtrim

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing and controlling this nondeterminism using a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction \cite{dpor} and Maximal Causality Reduction \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding \cite{chess-icb} allow bugs to be found sooner in a given exploration should they exist.

@@

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate OS classes at Carnegie Mellon, Berkeley, and the University of Chicago.
We show many advantages over both conventional stateless model checking and single-pass data-race analysis.
We find that many of the bugs found by the conventional approach can be found faster using subsets of preemption points,
and that data-race preemption points quickly expose many new bugs that prior model checkers could not find at all.

>>>>

% blah blah trite opening sentence
%As parallelism becomes ever more important for achieving high performance in modern-day programs,
%so too do advanced concurrency testing techniques become important for verifying the correctness of those programs.
Concurrency bugs are notoriously hard to find and reproduce because they only appear in specific thread interleavings, which arise at random during normal program execution.
{\em Stateless model checking} \cite{verisoft} offers a method for finding such bugs,
or verifying their absence,
%by systematically executing a program along as many distinct interleavings as possible,
by forcing a program to execute each distinct interleaving,
capturing
%and controlling
this nondeterminism in a finite state space.
Unfortunately, these state spaces explode exponentially in the size of the input program.
Reduction techniques such as Dynamic Partial Order Reduction \cite{dpor} and Maximal Causality Reduction \cite{mcr} expand the limits of feasible test completion,
and search ordering strategies such as Iterative Context Bounding \cite{chess-icb} allow bugs to be found sooner in a given space should they exist.

@@

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate OS classes at Carnegie Mellon, Berkeley, and the University of Chicago.
%We show many advantages over both conventional stateless model checking and single-pass data-race analysis.
We find many of the same bugs found by the conventional approach faster using subsets of preemption points,
and that data-race preemption points quickly expose many new bugs that prior model checkers could not find at all.

### 2015-11-18 pldi-16/overview.tex
trimtrimtrim

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{int x = 0; mutex\_t mx;}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{orange}{mutex\_lock(\&mx);}} & \\
	2 & \texttt{int tmp = x;} &\\
	3 & & \texttt{atomic\_xadd(\&x, 1);} \\
	4 & & \texttt{\hilight{olivegreen}{yield();}} \\
	5 & & \texttt{atomic\_xadd(\&x, 1);} \\
	6 & \texttt{x = tmp + 1;} & \\
	7 & \texttt{\hilight{commentblue}{mutex\_unlock(\&mx);}} & \\
	8 & \texttt{assert(x >= 1);}  & \\
	9 & & \texttt{assert(x >= 2);} \\
	%1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	%2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	%3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	%4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	%5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	%6 & & \texttt{publish(y);} \\
	%7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{Example program with a data-race bug. In this interleaving, the assertion on line 9 will fail. Two data-race preemptions are required to expose the bug: one just before thread 1's line 6, and one just before thread 2's line 9.}
\label{fig:example}
\end{figure}

@@

Data race analysis \cite{eraser} identifies pairs of unsynchronized memory accesses between threads.
Two instructions are said to race if
they both access the same memory address,
at least one is a write,
the threads do not hold the same lock,
and no synchronization enforces the order of the thread transitions.
Data races are not necessarily bugs, but represent suspicious violations of the locking discipline.
In Figure~\ref{fig:example}, lines 3 and 5 each race with 2, 6, and 8, and line 6 races with 9.

@@

For the rest of the paper, we will abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC), % (i.e., the approach of prior work),
Dynamic Partial Order Reduction (DPOR), and {\em state space estimate} (ETA).

SSS-MC indicates the approach of prior model checking tools:
the set of PPs is fixed in advance, and the tool commits to testing every interleaving available with those PPs.
%Reduction techniques are often used to skip equivalent interleavings \cite{dpor}, and search-ordering strategies
Many techniques exist to skip equivalent interleavings and order the search to uncover bugs faster \cite{dpor,demeter,chess-icb,gambit,smc-empirical-study},
but new PPs cannot be added, nor ineffective ones removed, by any dynamic analysis.

>>>>

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{int x = 0; mutex\_t mx;}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{orange}{mutex\_lock(\&mx);}} & \\
	2 & \texttt{int tmp = x;} &\\
	3 & & \texttt{atomic\_xadd(\&x, 1);} \\
	4 & & \texttt{\hilight{olivegreen}{yield();}} \\
	5 & & \texttt{atomic\_xadd(\&x, 1);} \\
	6 & \texttt{x = tmp + 1;} & \\
	7 & \texttt{\hilight{commentblue}{mutex\_unlock(\&mx);}} & \\
	8 & & \texttt{assert(x >= 2);} \\
	%1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	%2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	%3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	%4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	%5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	%6 & & \texttt{publish(y);} \\
	%7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{Example program with a data-race bug. In this interleaving, the assertion on line 8 will fail. Two data-race preemptions are required to expose the bug: one just before thread 1's line 6, and one just before thread 2's line 8.}
\label{fig:example}
\end{figure}

@@

Data race analysis \cite{eraser} identifies pairs of unsynchronized memory accesses between threads.
Two instructions are said to race if
they both access the same memory address,
at least one is a write,
the threads do not hold the same lock,
and no synchronization enforces the order of the thread transitions.
Data races are not necessarily bugs, but represent suspicious violations of the locking discipline.
In Figure~\ref{fig:example}, lines 3 and 5 each race with 2 and 6, and line 6 races with 8.

@@

For the rest of the paper, we will abbreviate {\em preemption point} (PP),
%For the remainder of the paper, we will abbreviate {\em preemption point} (PP),
{\em model checking} (MC),
{\em single-state-space model checking} (SSS-MC), % (i.e., the approach of prior work),
{\em Dynamic Partial Order Reduction} (DPOR), and {\em state space estimate} (ETA).

SSS-MC indicates the approach of prior tools:
the set of PPs is fixed in advance, and the tool commits to testing every interleaving available with those PPs.
%Reduction techniques are often used to skip equivalent interleavings \cite{dpor}, and search-ordering strategies
Many techniques can skip equivalent interleavings or order the search to uncover bugs faster \cite{dpor,demeter,chess-icb,gambit,smc-empirical-study},
but new PPs cannot be added, nor ineffective ones removed, by any dynamic analysis.
